Model-based pretreatment for regression with correlated variables.

### 2014-04-15 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@533 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

 	%plan
 	This paper will first present the marginal regression model and its properties before describing in Section 3 the random walk used to find the structure.
 	We will then look at some numerical results on simulated (Section \ref{sectionsimul}) and real industrial datasets (Section \ref{sectionrealcase}) before concluding and giving some perspectives in Section \ref{conclusion}.

\section{Model to select decorrelated covariates}
\subsection{A classical problem: correlations in regression}\label{sectionOLS}
For a model defined by 
	\begin{equation}
		\boldsymbol{Y}_{|\boldsymbol{X}}=\boldsymbol{X\beta} + \boldsymbol{\varepsilon}
	\end{equation}
	where $\boldsymbol{X}$ is the $n\times p$ matrix of the explicative variables, $\boldsymbol{Y}$ the response vector and $\boldsymbol{\varepsilon} \sim \mathcal{N}(\boldsymbol{0},\sigma_Y^2\boldsymbol{I_n})$ with $\boldsymbol{I_n}$ the $n$-sized identity matrix,
	we have the following Ordinary Least Squares (\textsc{OLS}) estimator:
	\begin{equation}
		\boldsymbol{\hat{\beta}}=\left(\boldsymbol{X}'\boldsymbol{X} \right) ^{-1}\boldsymbol{X}'\boldsymbol{Y}
	\end{equation}
	with variance matrix
	\begin{equation}
		\operatorname{Var}(\boldsymbol{\hat{\beta}})=\sigma_Y^2\left(\boldsymbol{X}'\boldsymbol{X} \right) ^{-1} \label{eqOLS}
	\end{equation}
	and without any bias.
	When correlations between covariates are strong, the matrix to invert is ill-conditioned and the variance increases, giving unstable and unusable estimator \cite{hoerl1970ridge}.
	Another problem is that matricial inversion requires $n\geq p$. 
\subsection{Our proposal: modelisation of the correlations}
Let now $\boldsymbol{X} \in \mathcal{R}^{n\times p}$ be a set of $p$ covariates.
In the following, we note $\boldsymbol{X^j}$ the $j^{th}$ column of $\boldsymbol{X}$.

We propose to explicitly define a family of $p_r$ internal regressions between covariates with $I_r$ the set of indices of endogenous variables in $\boldsymbol{X}$ (explained ones). We note $\boldsymbol{X}_r=\boldsymbol{X}^{I_r}$ the sub-matrix of the redundant covariates (explained by others) and  $\boldsymbol{X}_f$ the remaining submatrix of free covariates (independent explaining covariates).
We focus here on the expression of $P(\boldsymbol{X}_r|\boldsymbol{X}_f)$.

@@

We make the hypothesis of the uncrossing rule $I_f\cap I_r=\emptyset$ {\it i.e.} endogenous variables don't explain other covariates, thus $\boldsymbol{X}_f$ and $\boldsymbol{X}_f$ give a partition of $\boldsymbol{X}$, so $\boldsymbol{X}_f=\boldsymbol{X}\setminus \boldsymbol{X}_r$.
The discrete structure $S$ is identifiable because we can't permute some regressions in (\ref{SR}) and obtain the same joint distribution $P(\boldsymbol{X},\boldsymbol{Y})$, the residuals ($\boldsymbol{\varepsilon}_j$) would not stay Gaussian (see the appendices).

@@

	  We make the choice to distinguish the response variable from the other endogenous variables (that are on the left of a regression). Thus we have one regression on our response variable ($P(\boldsymbol{Y}|\boldsymbol{X})$ and a system of sub-regressions (without the response variable: $P(\boldsymbol{X}_r|\boldsymbol{X}_f,S)$). Then we consider correlations between the explicative covariates of the main regression, not between the residuals.

>>>>

 	%plan
 	This paper will first present the linear modelisation of the correlations and the by-product marginal regression model before describing in Section 3 the random walk used to find the structure.
 	We will then look at some numerical results on simulated (Section \ref{sectionsimul}) and real industrial datasets (Section \ref{sectionrealcase}) before concluding and giving some perspectives in Section \ref{conclusion}.

\section{Model to select decorrelated covariates}
\subsection{A classical problem: correlations in regression}\label{sectionOLS}
For a model defined by (\ref{regressionsimple}), we have the following Ordinary Least Squares (\textsc{OLS}) estimator:
	\begin{equation}
		\boldsymbol{\hat{\beta}}=\left(\boldsymbol{X}'\boldsymbol{X} \right) ^{-1}\boldsymbol{X}'\boldsymbol{Y}
	\end{equation}
	with variance matrix
	\begin{equation}
		\operatorname{Var}(\boldsymbol{\hat{\beta}})=\sigma_Y^2\left(\boldsymbol{X}'\boldsymbol{X} \right) ^{-1} \label{eqOLS}
	\end{equation}
	and without any bias.
	When correlations between covariates are strong, the matrix to invert is ill-conditioned and the variance increases, giving unstable and unusable estimator \cite{hoerl1970ridge}.
	Another problem is that matrix inversion requires $n\geq p$. 
\subsection{Our proposal: modelisation of the correlations}
Let now $\boldsymbol{X} \in \mathcal{R}^{n\times p}$ be a set of $p$ covariates.
In the following, we note $\boldsymbol{X^j}$ the $j^{th}$ column of $\boldsymbol{X}$.

We make the hypothesis that correlations in $\boldsymbol{X}$ are generated by a family of $p_r$ internal regressions between covariates with $I_r$ the set of indices of endogenous variables in $\boldsymbol{X}$ (explained ones). We note $\boldsymbol{X}_r=\boldsymbol{X}^{I_r}$ the sub-matrix of the redundant covariates (explained by others) and  $\boldsymbol{X}_f$ the remaining submatrix of free covariates (independent explaining covariates).
We focus here on the expression of $P(\boldsymbol{X}_r|\boldsymbol{X}_f)$.

@@

Now we have an explicit structure of correlations and we know which covariates will cause ill-conditioning. The explicit form also gives the ability to understand phenomenon inside the dataset and to see why some covariates are not selected by methods like the LASSO when all $\boldsymbol{X}$ is provided (redundent variables).

We make the hypothesis of the uncrossing rule $I_f\cap I_r=\emptyset$ {\it i.e.} endogenous variables don't explain other covariates, thus $\boldsymbol{X}_r$ and $\boldsymbol{X}_f$ give a partition of $\boldsymbol{X}$, so $\boldsymbol{X}_f=\boldsymbol{X}\setminus \boldsymbol{X}_r$. This hypothesis ensures that $S$ contains no cycle and is straightforward readable (no need to order $I_r$). This hypothesis is not so restrictive because cyclic structures have no sense and any non-cyclic structure can be associated with a structure that verifies the uncrossing constraint by just replacing covariates in $\boldsymbol{X}_r$ by their expression in $S$ when on the right of a sub-regression.

The discrete structure $S$ is identifiable because we can't permute some regressions in (\ref{SR}) and obtain the same joint distribution $P(\boldsymbol{X},\boldsymbol{Y})$, the residuals ($\boldsymbol{\varepsilon}_j$) would not stay Gaussian (see the appendices).

@@

	  We make the choice to distinguish the response variable from the other endogenous variables (that are on the left of a regression). Thus we have one regression on our response variable ($P(\boldsymbol{Y}|\boldsymbol{X})$ and a system of sub-regressions (without the response variable: $P(\boldsymbol{X}_r|\boldsymbol{X}_f,S)$). Then we consider correlations between the explicative covariates of the main regression, not between the residuals. We see that the $S$ does not depend on $Y$ so it can be learnt independently, even with a larger dataset (if missing values in $Y$).
